---
title: "try_Seurat"
author: "xzk"
date: '2022-05-13'
output: html_document
---

# integrate mouse and human T cells with Seurat_v3

## load packages
```{r}
suppressPackageStartupMessages({
  library(SingleCellExperiment)
  library(scMerge)
  library(zellkonverter)
  library(Seurat)
  library(scater)
})
```

## load mouse and human T cells data

```{r}
setwd('D:/xuzk')
LCA_T <- readH5AD('data/raw_combined_LCA_T_n.h5ad')
assays(LCA_T)$counts=assays(LCA_T)$X
assays(LCA_T) <- assays(LCA_T)[-1]
LCA_T <- logNormCounts(LCA_T)
LCA_T <- as.Seurat(LCA_T)
```

## split the dataset into a list of two seurat objects (mouse and human)
```{r}
LCA_T.list <- SplitObject(LCA_T, split.by = "species")
rm(LCA_T)
```

## normalization, HVG and integration

```{r}
# normalize and identify variable features for each dataset independently
LCA_T.list <- lapply(X = LCA_T.list, FUN = function(x) {
  x <- NormalizeData(x)
  x <- FindVariableFeatures(x, selection.method = "vst", nfeatures = 2000)
})

# select features that are repeatedly variable across datasets for integration
features <- SelectIntegrationFeatures(object.list = LCA_T.list)


# perform integration
LCA_T.anchors <- FindIntegrationAnchors(object.list = LCA_T.list, 
                                         anchor.features = features)
# this command creates an 'integrated' data assay
LCA_T.combined <- IntegrateData(anchorset = LCA_T.anchors)
```

# downstream analysis on the corrected data 
```{r pressure, echo=FALSE}
DefaultAssay(LCA_T.combined) <- "integrated"
# Run the standard workflow for visualization and clustering
LCA_T.combined <- ScaleData(LCA_T.combined,vars.to.regress = c("total_counts", "pct_counts_mt"), verbose = FALSE)
LCA_T.combined <- RunPCA(LCA_T.combined, npcs = 30, verbose = FALSE)

ElbowPlot(LCA_T.combined )

LCA_T.combined <- RunUMAP(LCA_T.combined, reduction = "pca", dims = 1:8)
LCA_T.combined <- FindNeighbors(LCA_T.combined, reduction = "pca", dims = 1:30)
LCA_T.combined <- FindClusters(LCA_T.combined, resolution = 0.5)
```

# Visualization
```{r, echo=TRUE}
DimPlot(LCA_T.combined, reduction = "umap", group.by = "species")

DimPlot(LCA_T.combined, reduction = "umap", group.by = "label_subtype", cols='polychrome')

# label_subtype-species
DimPlot(LCA_T.combined, reduction = "umap", group.by = "label_subtype", split.by = 'species',cols='polychrome')
DimPlot(LCA_T.combined, reduction = "umap", group.by = "label_subtype_hres", split.by = 'species',cols='polychrome')


# cluster
DimPlot(LCA_T.combined, reduction = "umap", label = TRUE, repel = TRUE, cols='alphabet')
DimPlot(LCA_T.combined, reduction = "umap", group.by = "seurat_clusters", split.by = 'species', cols='alphabet')



```

# DEG
# DEGs between CD8+ effector T cells
```{r}
#colnames(LCA_T.combined[[]])
#Idents(LCA_T.combined) <- 'label_subtype_hres'
LCA_T.CD8eff <- subset(LCA_T.combined,idents=c("h-CD8+ cytotoxic 1","h-CD8+ cytotoxic 2 (IL7R, gamma-delta)","h-CD8+ cytotoxic & exhausted","h-CD8+ cytotoxic & exhausted (gamma-delta)","m-CD8+ cytotoxic T"))
Teff_markers <- FindMarkers(LCA_T.CD8eff, ident.1 = "m-CD8+ cytotoxic T", group.by='label_subtype_hres',logfc.threshold = 0 )
```
# DEGs between Treg,Tfh cells
```{r}
#colnames(LCA_T.combined[[]])
#Idents(LCA_T.combined) <- 'label_subtype_hres'
LCA_T.Tregfh <- subset(LCA_T.combined,idents=c('h-CD4+ Treg','h-CD4+ Tfh','m-CD4+ exhausted Tfh&Treg'))
Tregfh_markers <- FindMarkers(LCA_T.Tregfh, ident.1 = "m-CD4+ exhausted Tfh&Treg", group.by='label_subtype_hres',logfc.threshold = 0 )
```

# DEGs between CD4+ T cells
```{r}
#colnames(LCA_T.combined[[]])
#Idents(LCA_T.combined) <- 'label_subtype_hres'
LCA_T.CD4 <- subset(LCA_T.combined,idents=c('h-CD4+ naive T','h-CD4+ undefined T','h-Th17-like','h-Th2-like','m-CD4+ naive T','m-Th9-like'))
TCD4_markers <- FindMarkers(LCA_T.CD4, ident.1 = "h-CD4+ undefined T", group.by='label_subtype_hres',logfc.threshold = 0 )
```

# DEG
# DEG between 2 subtypes
```{r}
T_markers <- FindMarkers(LCA_T.combined, ident.1 = "m-CD8+ cytotoxic T", ident.2='h-CD8+ cytotoxic 2 (IL7R, gamma-delta)',group.by='label_subtype_hres',logfc.threshold = 1 )
setwd("D:/xuzk")
write.table(T_markers,'results/DEG-m-ct_h-ct2.txt')
```



```{r}
setwd("D:/xuzk")
#saveRDS(LCA_T.combined, file = "outputs/LCA_T.rds")
LCA_T.combined <- readRDS("outputs/LCA_T.rds")
```

